---
title: "abandoned_cropland_Brazil"
author: "Nick McManus"
date: '2022-10-06'
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: no
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)  #polygons
library(raster) #raster package
library(terra)  #newer/faster/better raster package
library(rnaturalearth)  #administrative boundaries data
library(tidyverse) #always
```

## Overview
This analysis serves to find areas of projected cropland abandonment in Brazil between the years 2020 to 2050. Land use and land cover (LULC) data from Chen et al. 2022. This source includes both projected LULC under different shared socioeconomic pathways (SSPs) as well as a baseline LULC data from 2015. The outputs from this code, rasters containing every pixel of projected abandonment between 2020-2050 for a given SSP, will be utilized for further analysis into conservation optimization. More information on the source data can be found in the project README.md file. 

Source: Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6


## Read in data

First read in projected LULC from Chen et al. for every five years between 2020-2050. There are LULC layers for each of the five different SSPs.

```{r}
# read in "baseline" 2015 LULC
PFT_2015_global <- rast("data/src/global_PFT_2015.tif")


# read in projected SSP1 LULC for 2020 - 2050 -------------------------------------------
ssp1_2020_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2020.tif")   #2020
ssp1_2025_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2025.tif")   #2025
ssp1_2030_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2030.tif")   #2030
ssp1_2035_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2035.tif")   #2035
ssp1_2040_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2040.tif")   #2040
ssp1_2045_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2045.tif")   #2045
ssp1_2050_global <- rast("data/src/SSP1_RCP26/global_PFT_SSP1_RCP26_2050.tif")   #2050


# read in projected SSP2 LULC for 2020 - 2050 -------------------------------------------
ssp2_2020_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2020.tif") 
ssp2_2025_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2025.tif")
ssp2_2030_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2030.tif") 
ssp2_2035_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2035.tif")
ssp2_2040_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2040.tif")
ssp2_2045_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2045.tif")
ssp2_2050_global <- rast("data/src/SSP2_RCP45/global_PFT_SSP2_RCP45_2050.tif")


# read in projected SSP3 LULC for 2020 - 2050 -------------------------------------------
ssp3_2020_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2020.tif")  
ssp3_2025_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2025.tif")  
ssp3_2030_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2030.tif")  
ssp3_2035_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2035.tif")  
ssp3_2040_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2040.tif")  
ssp3_2045_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2045.tif")  
ssp3_2050_global <- rast("data/src/SSP3_RCP70/global_PFT_SSP3_RCP70_2050.tif")  


# read in projected SSP4 LULC for 2020 - 2050 -------------------------------------------
ssp4_2020_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2020.tif") 
ssp4_2025_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2025.tif") 
ssp4_2030_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2030.tif") 
ssp4_2035_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2035.tif") 
ssp4_2040_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2040.tif") 
ssp4_2045_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2045.tif")
ssp4_2050_global <- rast("data/src/SSP4_RCP60/global_PFT_SSP4_RCP60_2050.tif") 


# read in projected SSP5 LULC for 2020 - 2050 -------------------------------------------
ssp5_2020_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2020.tif") 
ssp5_2025_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2025.tif")
ssp5_2030_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2030.tif")
ssp5_2035_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2035.tif")
ssp5_2040_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2040.tif")
ssp5_2045_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2045.tif")
ssp5_2050_global <- rast("data/src/SSP5_RCP85/global_PFT_SSP5_RCP85_2050.tif")
```


Brazil is the ROI for this analysis. Read in shape file for Brazil to later clip LULC layers. 
```{r}
# get brazil from rnaturalearth data
brazil <- ne_countries(
  scale = "medium",
  country = "Brazil",
  returnclass = "sf") %>% 
  dplyr::select(sovereignt)

# plot to check
ggplot() +
  geom_sf(data = brazil)
```


## Clipping and Masking 

Transform Brazil shapefile to match Chen LULC raster coordinate reference system. This is faster then re-projected a global raster to match Brazil's crs.
```{r}
# transform brazil to chen raster
brazil_trans <- brazil %>% 
  st_transform(crs = crs(PFT_2015_global)) %>% 
  vect() # make object a vector, b/c vects more agreeable to rasts
```

Now crop and mask each LULC raster to our `brazil_trans` vector. Then, project raster back to our desired crs.
- Note: the `terra::crop()` function that clips rasters is not to be confused with "crops" used in later variable names to indicate rasters reclassified to isolate cropland. 
```{r}
# crop baseline 2015 data
PFT_2015_brazil <- PFT_2015_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP1 layers --------------------------
ssp1_2020_brazil <- ssp1_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2025_brazil <- ssp1_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2030_brazil <- ssp1_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2035_brazil <- ssp1_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2040_brazil <- ssp1_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2045_brazil <- ssp1_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2050_brazil <- ssp1_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP2 layers --------------------------
ssp2_2020_brazil <- ssp2_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2025_brazil <- ssp2_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2030_brazil <- ssp2_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2035_brazil <- ssp2_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2040_brazil <- ssp2_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2045_brazil <- ssp2_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2050_brazil <- ssp2_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP3 layers --------------------------
ssp3_2020_brazil <- ssp3_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2025_brazil <- ssp3_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2030_brazil <- ssp3_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2035_brazil <- ssp3_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2040_brazil <- ssp3_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2045_brazil <- ssp3_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2050_brazil <- ssp3_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP4 layers ---------------------------
ssp4_2020_brazil <- ssp4_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2025_brazil <- ssp4_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2030_brazil <- ssp4_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2035_brazil <- ssp4_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2040_brazil <- ssp4_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2045_brazil <- ssp4_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2050_brazil <- ssp4_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP5 layers --------------------------
ssp5_2020_brazil <- ssp5_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2025_brazil <- ssp5_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2030_brazil <- ssp5_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2035_brazil <- ssp5_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2040_brazil <- ssp5_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2045_brazil <- ssp5_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2050_brazil <- ssp5_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))
```


## Reclassify

First, reclassify each LULC layer to isolate pixels of cropland (value 18) from the raster. 
- Note: this method for reclassifying does not use the `terra::classify()` function. The same step could be done with `classify()`, but would require creating a matrix reassigning each of the 20 values for the raster. Because we're only interested in keeping one value from the raster, this was a quick work-around. 
- NOTE: this method for reclassifying and calculating abandonment was actually longer than necessary! See workflow in "abandoned-cropland-global" markdown to see how the process could be expedited by reclassifying and subtracting rasters just one time. This analysis has already been run and the layers exported, so there was no immediate need to re-write the inefficient code.

```{r}
# baseline 2015 data
crops_2015_brazil <- PFT_2015_brazil 
  crops_2015_brazil[crops_2015_brazil != 18] <- 0 # if value isn't 18, make 0
  crops_2015_brazil[crops_2015_brazil == 18] <- 1 # if value is 18, make 1
  
  
# SSP1 layers for 2020 - 2050 ---------------------------
crops_ssp1_2020 <- ssp1_2020_brazil
  crops_ssp1_2020[crops_ssp1_2020 != 18] <- 0
  crops_ssp1_2020[crops_ssp1_2020 == 18] <- 1

crops_ssp1_2025 <- ssp1_2025_brazil
  crops_ssp1_2025[crops_ssp1_2025 != 18] <- 0
  crops_ssp1_2025[crops_ssp1_2025 == 18] <- 1
  
crops_ssp1_2030 <- ssp1_2030_brazil   
  crops_ssp1_2030[crops_ssp1_2030 != 18] <- 0
  crops_ssp1_2030[crops_ssp1_2030 == 18] <- 1
  
crops_ssp1_2035 <- ssp1_2035_brazil
  crops_ssp1_2035[crops_ssp1_2035 != 18] <- 0
  crops_ssp1_2035[crops_ssp1_2035 == 18] <- 1  
  
crops_ssp1_2040 <- ssp1_2040_brazil
  crops_ssp1_2040[crops_ssp1_2040 != 18] <- 0
  crops_ssp1_2040[crops_ssp1_2040 == 18] <- 1
  
crops_ssp1_2045 <- ssp1_2045_brazil
  crops_ssp1_2045[crops_ssp1_2045 != 18] <- 0
  crops_ssp1_2045[crops_ssp1_2045 == 18] <- 1
  
crops_ssp1_2050 <- ssp1_2050_brazil
  crops_ssp1_2050[crops_ssp1_2050 != 18] <- 0
  crops_ssp1_2050[crops_ssp1_2050 == 18] <- 1
  

# SSP2 layers for 2020 - 2050 ---------------------------
crops_ssp2_2020 <- ssp2_2020_brazil
  crops_ssp2_2020[crops_ssp2_2020 != 18] <- 0
  crops_ssp2_2020[crops_ssp2_2020 == 18] <- 1
  
crops_ssp2_2025 <- ssp2_2025_brazil
  crops_ssp2_2025[crops_ssp2_2025 != 18] <- 0
  crops_ssp2_2025[crops_ssp2_2025 == 18] <- 1

crops_ssp2_2030 <- ssp2_2030_brazil   
  crops_ssp2_2030[crops_ssp2_2030 != 18] <- 0
  crops_ssp2_2030[crops_ssp2_2030 == 18] <- 1
  
crops_ssp2_2035 <- ssp2_2035_brazil
  crops_ssp2_2035[crops_ssp2_2035 != 18] <- 0
  crops_ssp2_2035[crops_ssp2_2035 == 18] <- 1
  
crops_ssp2_2040 <- ssp2_2040_brazil
  crops_ssp2_2040[crops_ssp2_2040 != 18] <- 0
  crops_ssp2_2040[crops_ssp2_2040 == 18] <- 1
  
crops_ssp2_2045 <- ssp2_2045_brazil
  crops_ssp2_2045[crops_ssp2_2045 != 18] <- 0
  crops_ssp2_2045[crops_ssp2_2045 == 18] <- 1
  
crops_ssp2_2050 <- ssp2_2050_brazil
  crops_ssp2_2050[crops_ssp2_2050 != 18] <- 0
  crops_ssp2_2050[crops_ssp2_2050 == 18] <- 1

  
# SSP3 layers for 2020 - 2050 ---------------------------
crops_ssp3_2020 <- ssp3_2020_brazil
  crops_ssp3_2020[crops_ssp3_2020 != 18] <- 0
  crops_ssp3_2020[crops_ssp3_2020 == 18] <- 1
  
crops_ssp3_2025 <- ssp3_2025_brazil
  crops_ssp3_2025[crops_ssp3_2025 != 18] <- 0
  crops_ssp3_2025[crops_ssp3_2025 == 18] <- 1

crops_ssp3_2030 <- ssp3_2030_brazil   
  crops_ssp3_2030[crops_ssp3_2030 != 18] <- 0
  crops_ssp3_2030[crops_ssp3_2030 == 18] <- 1
  
crops_ssp3_2035 <- ssp3_2035_brazil
  crops_ssp3_2035[crops_ssp3_2035 != 18] <- 0
  crops_ssp3_2035[crops_ssp3_2035 == 18] <- 1
  
crops_ssp3_2040 <- ssp3_2040_brazil
  crops_ssp3_2040[crops_ssp3_2040 != 18] <- 0
  crops_ssp3_2040[crops_ssp3_2040 == 18] <- 1
  
crops_ssp3_2045 <- ssp3_2045_brazil
  crops_ssp3_2045[crops_ssp3_2045 != 18] <- 0
  crops_ssp3_2045[crops_ssp3_2045 == 18] <- 1
  
crops_ssp3_2050 <- ssp3_2050_brazil
  crops_ssp3_2050[crops_ssp3_2050 != 18] <- 0
  crops_ssp3_2050[crops_ssp3_2050 == 18] <- 1
  

# SSP4 layers for 2020 - 2050 ---------------------------
crops_ssp4_2020 <- ssp4_2020_brazil
  crops_ssp4_2020[crops_ssp4_2020 != 18] <- 0
  crops_ssp4_2020[crops_ssp4_2020 == 18] <- 1
  
crops_ssp4_2025 <- ssp4_2025_brazil
  crops_ssp4_2025[crops_ssp4_2025 != 18] <- 0
  crops_ssp4_2025[crops_ssp4_2025 == 18] <- 1

crops_ssp4_2030 <- ssp4_2030_brazil   
  crops_ssp4_2030[crops_ssp4_2030 != 18] <- 0
  crops_ssp4_2030[crops_ssp4_2030 == 18] <- 1
  
crops_ssp4_2035 <- ssp4_2035_brazil
  crops_ssp4_2035[crops_ssp4_2035 != 18] <- 0
  crops_ssp4_2035[crops_ssp4_2035 == 18] <- 1
  
crops_ssp4_2040 <- ssp4_2040_brazil
  crops_ssp4_2040[crops_ssp4_2040 != 18] <- 0
  crops_ssp4_2040[crops_ssp4_2040 == 18] <- 1
  
crops_ssp4_2045 <- ssp4_2045_brazil
  crops_ssp4_2045[crops_ssp4_2045 != 18] <- 0
  crops_ssp4_2045[crops_ssp4_2045 == 18] <- 1
  
crops_ssp4_2050 <- ssp4_2050_brazil
  crops_ssp4_2050[crops_ssp4_2050 != 18] <- 0
  crops_ssp4_2050[crops_ssp4_2050 == 18] <- 1
  
  
# SSP5 layers for 2020 - 2050 ---------------------------
crops_ssp5_2020 <- ssp5_2020_brazil
  crops_ssp5_2020[crops_ssp5_2020 != 18] <- 0
  crops_ssp5_2020[crops_ssp5_2020 == 18] <- 1
  
crops_ssp5_2025 <- ssp5_2025_brazil
  crops_ssp5_2025[crops_ssp5_2025 != 18] <- 0
  crops_ssp5_2025[crops_ssp5_2025 == 18] <- 1

crops_ssp5_2030 <- ssp5_2030_brazil   
  crops_ssp5_2030[crops_ssp5_2030 != 18] <- 0
  crops_ssp5_2030[crops_ssp5_2030 == 18] <- 1
  
crops_ssp5_2035 <- ssp5_2035_brazil
  crops_ssp5_2035[crops_ssp5_2035 != 18] <- 0
  crops_ssp5_2035[crops_ssp5_2035 == 18] <- 1
  
crops_ssp5_2040 <- ssp5_2040_brazil
  crops_ssp5_2040[crops_ssp5_2040 != 18] <- 0
  crops_ssp5_2040[crops_ssp5_2040 == 18] <- 1
  
crops_ssp5_2045 <- ssp5_2045_brazil
  crops_ssp5_2045[crops_ssp5_2045 != 18] <- 0
  crops_ssp5_2045[crops_ssp5_2045 == 18] <- 1
  
crops_ssp5_2050 <- ssp5_2050_brazil
  crops_ssp5_2050[crops_ssp5_2050 != 18] <- 0
  crops_ssp5_2050[crops_ssp5_2050 == 18] <- 1

```


Next, reclassify layers each projected LULC layer to isolate urban pixels (value 19). Use for these will be explained in the following section.
```{r}
# SSP1 layers for 2020 - 2050 ------------------------
urban_ssp1_2020 <- ssp1_2020_brazil
  urban_ssp1_2020[urban_ssp1_2020 != 19] <- 0   # if value isn't 19, make 0
  urban_ssp1_2020[urban_ssp1_2020 == 19] <- 10  # if value is 19, make 10

urban_ssp1_2025 <- ssp1_2025_brazil
  urban_ssp1_2025[urban_ssp1_2025 != 19] <- 0
  urban_ssp1_2025[urban_ssp1_2025 == 19] <- 10
  
urban_ssp1_2030 <- ssp1_2030_brazil
  urban_ssp1_2030[urban_ssp1_2030 != 19] <- 0
  urban_ssp1_2030[urban_ssp1_2030 == 19] <- 10
  
urban_ssp1_2035 <- ssp1_2035_brazil
  urban_ssp1_2035[urban_ssp1_2035 != 19] <- 0
  urban_ssp1_2035[urban_ssp1_2035 == 19] <- 10
  
urban_ssp1_2040 <- ssp1_2040_brazil
  urban_ssp1_2040[urban_ssp1_2040 != 19] <- 0
  urban_ssp1_2040[urban_ssp1_2040 == 19] <- 10
  
urban_ssp1_2045 <- ssp1_2045_brazil
  urban_ssp1_2045[urban_ssp1_2045 != 19] <- 0
  urban_ssp1_2045[urban_ssp1_2045 == 19] <- 10
  
urban_ssp1_2050 <- ssp1_2050_brazil
  urban_ssp1_2050[urban_ssp1_2050 != 19] <- 0
  urban_ssp1_2050[urban_ssp1_2050 == 19] <- 10
  

# SSP2 layers for 2020 - 2050 ------------------------
urban_ssp2_2020 <- ssp2_2020_brazil
  urban_ssp2_2020[urban_ssp2_2020 != 19] <- 0  
  urban_ssp2_2020[urban_ssp2_2020 == 19] <- 10
  
urban_ssp2_2025 <- ssp2_2025_brazil
  urban_ssp2_2025[urban_ssp2_2025 != 19] <- 0  
  urban_ssp2_2025[urban_ssp2_2025 == 19] <- 10

urban_ssp2_2030 <- ssp2_2030_brazil   
  urban_ssp2_2030[urban_ssp2_2030 != 19] <- 0
  urban_ssp2_2030[urban_ssp2_2030 == 19] <- 10
  
urban_ssp2_2035 <- ssp2_2035_brazil
  urban_ssp2_2035[urban_ssp2_2035 != 19] <- 0  
  urban_ssp2_2035[urban_ssp2_2035 == 19] <- 10
  
urban_ssp2_2040 <- ssp2_2040_brazil
  urban_ssp2_2040[urban_ssp2_2040 != 19] <- 0
  urban_ssp2_2040[urban_ssp2_2040 == 19] <- 10
  
urban_ssp2_2045 <- ssp2_2045_brazil
  urban_ssp2_2045[urban_ssp2_2045 != 19] <- 0  
  urban_ssp2_2045[urban_ssp2_2045 == 19] <- 10
  
urban_ssp2_2050 <- ssp2_2050_brazil
  urban_ssp2_2050[urban_ssp2_2050 != 19] <- 0
  urban_ssp2_2050[urban_ssp2_2050 == 19] <- 10

  
# SSP3 layers for 2020 - 2050 ------------------------
urban_ssp3_2020 <- ssp3_2020_brazil
  urban_ssp3_2020[urban_ssp3_2020 != 19] <- 0  
  urban_ssp3_2020[urban_ssp3_2020 == 19] <- 10
  
urban_ssp3_2025 <- ssp3_2025_brazil
  urban_ssp3_2025[urban_ssp3_2025 != 19] <- 0  
  urban_ssp3_2025[urban_ssp3_2025 == 19] <- 10

urban_ssp3_2030 <- ssp3_2030_brazil   
  urban_ssp3_2030[urban_ssp3_2030 != 19] <- 0
  urban_ssp3_2030[urban_ssp3_2030 == 19] <- 10
  
urban_ssp3_2035 <- ssp3_2035_brazil
  urban_ssp3_2035[urban_ssp3_2035 != 19] <- 0  
  urban_ssp3_2035[urban_ssp3_2035 == 19] <- 10
  
urban_ssp3_2040 <- ssp3_2040_brazil
  urban_ssp3_2040[urban_ssp3_2040 != 19] <- 0
  urban_ssp3_2040[urban_ssp3_2040 == 19] <- 10
  
urban_ssp3_2045 <- ssp3_2045_brazil
  urban_ssp3_2045[urban_ssp3_2045 != 19] <- 0  
  urban_ssp3_2045[urban_ssp3_2045 == 19] <- 10
  
urban_ssp3_2050 <- ssp3_2050_brazil
  urban_ssp3_2050[urban_ssp3_2050 != 19] <- 0
  urban_ssp3_2050[urban_ssp3_2050 == 19] <- 10
  

# SSP4 layers for 2020 - 2050 ------------------------
urban_ssp4_2020 <- ssp4_2020_brazil
  urban_ssp4_2020[urban_ssp4_2020 != 19] <- 0  
  urban_ssp4_2020[urban_ssp4_2020 == 19] <- 10
  
urban_ssp4_2025 <- ssp4_2025_brazil
  urban_ssp4_2025[urban_ssp4_2025 != 19] <- 0  
  urban_ssp4_2025[urban_ssp4_2025 == 19] <- 10

urban_ssp4_2030 <- ssp4_2030_brazil   
  urban_ssp4_2030[urban_ssp4_2030 != 19] <- 0
  urban_ssp4_2030[urban_ssp4_2030 == 19] <- 10
  
urban_ssp4_2035 <- ssp4_2035_brazil
  urban_ssp4_2035[urban_ssp4_2035 != 19] <- 0  
  urban_ssp4_2035[urban_ssp4_2035 == 19] <- 10
  
urban_ssp4_2040 <- ssp4_2040_brazil
  urban_ssp4_2040[urban_ssp4_2040 != 19] <- 0
  urban_ssp4_2040[urban_ssp4_2040 == 19] <- 10
  
urban_ssp4_2045 <- ssp4_2045_brazil
  urban_ssp4_2045[urban_ssp4_2045 != 19] <- 0  
  urban_ssp4_2045[urban_ssp4_2045 == 19] <- 10
  
urban_ssp4_2050 <- ssp4_2050_brazil
  urban_ssp4_2050[urban_ssp4_2050 != 19] <- 0
  urban_ssp4_2050[urban_ssp4_2050 == 19] <- 10
  
  
# SSP5 layers for 2020 - 2050 ------------------------
urban_ssp5_2020 <- ssp5_2020_brazil
  urban_ssp5_2020[urban_ssp5_2020 != 19] <- 0  
  urban_ssp5_2020[urban_ssp5_2020 == 19] <- 10
  
urban_ssp5_2025 <- ssp5_2025_brazil
  urban_ssp5_2025[urban_ssp5_2025 != 19] <- 0  
  urban_ssp5_2025[urban_ssp5_2025 == 19] <- 10

urban_ssp5_2030 <- ssp5_2030_brazil   
  urban_ssp5_2030[urban_ssp5_2030 != 19] <- 0
  urban_ssp5_2030[urban_ssp5_2030 == 19] <- 10
  
urban_ssp5_2035 <- ssp5_2035_brazil
  urban_ssp5_2035[urban_ssp5_2035 != 19] <- 0  
  urban_ssp5_2035[urban_ssp5_2035 == 19] <- 10
  
urban_ssp5_2040 <- ssp5_2040_brazil
  urban_ssp5_2040[urban_ssp5_2040 != 19] <- 0
  urban_ssp5_2040[urban_ssp5_2040 == 19] <- 10
  
urban_ssp5_2045 <- ssp5_2045_brazil
  urban_ssp5_2045[urban_ssp5_2045 != 19] <- 0  
  urban_ssp5_2045[urban_ssp5_2045 == 19] <- 10
  
urban_ssp5_2050 <- ssp5_2050_brazil
  urban_ssp5_2050[urban_ssp5_2050 != 19] <- 0
  urban_ssp5_2050[urban_ssp5_2050 == 19] <- 10
```


## Calculate abandonment

Now that we've isolated Brazilian cropland, we want to determine where current cropland is projected to become "abandoned" in the future. This will be done with a simple raster calculation, subtracting the future cropland raster from the current cropland raster. Values of 0 indicate land that either remained cropland (1-1) or remained non-cropland (0-0). Values of -1 indicate new future cropland (0-1), while values of 1 indicate abandoned cropland (1-0), our pixels of interest! 

This calculation will be repeated for each projected raster (2020-2050) against the baseline raster (2015). 

```{r}
# SSP1 abandonment 2020-2050 ----------------------------
ab_ssp1_2020 <- crops_2015_brazil - crops_ssp1_2020
ab_ssp1_2025 <- crops_2015_brazil - crops_ssp1_2025
ab_ssp1_2030 <- crops_2015_brazil - crops_ssp1_2030
ab_ssp1_2035 <- crops_2015_brazil - crops_ssp1_2035
ab_ssp1_2040 <- crops_2015_brazil - crops_ssp1_2040
ab_ssp1_2045 <- crops_2015_brazil - crops_ssp1_2045
ab_ssp1_2050 <- crops_2015_brazil - crops_ssp1_2050


# SSP2 abandonment 2020-2050 ----------------------------
ab_ssp2_2020 <- crops_2015_brazil - crops_ssp2_2020
ab_ssp2_2025 <- crops_2015_brazil - crops_ssp2_2025
ab_ssp2_2030 <- crops_2015_brazil - crops_ssp2_2030
ab_ssp2_2035 <- crops_2015_brazil - crops_ssp2_2035
ab_ssp2_2040 <- crops_2015_brazil - crops_ssp2_2040
ab_ssp2_2045 <- crops_2015_brazil - crops_ssp2_2045
ab_ssp2_2050 <- crops_2015_brazil - crops_ssp2_2050


# SSP3 abandonment 2020-2050 ----------------------------
ab_ssp3_2020 <- crops_2015_brazil - crops_ssp3_2020
ab_ssp3_2025 <- crops_2015_brazil - crops_ssp3_2025
ab_ssp3_2030 <- crops_2015_brazil - crops_ssp3_2030
ab_ssp3_2035 <- crops_2015_brazil - crops_ssp3_2035
ab_ssp3_2040 <- crops_2015_brazil - crops_ssp3_2040
ab_ssp3_2045 <- crops_2015_brazil - crops_ssp3_2045
ab_ssp3_2050 <- crops_2015_brazil - crops_ssp3_2050


# SSP4 abandonment 2020-2050 ----------------------------
ab_ssp4_2020 <- crops_2015_brazil - crops_ssp4_2020
ab_ssp4_2025 <- crops_2015_brazil - crops_ssp4_2025
ab_ssp4_2030 <- crops_2015_brazil - crops_ssp4_2030
ab_ssp4_2035 <- crops_2015_brazil - crops_ssp4_2035
ab_ssp4_2040 <- crops_2015_brazil - crops_ssp4_2040
ab_ssp4_2045 <- crops_2015_brazil - crops_ssp4_2045
ab_ssp4_2050 <- crops_2015_brazil - crops_ssp4_2050


# SSP5 abandonment 2020-2050 ----------------------------
ab_ssp5_2020 <- crops_2015_brazil - crops_ssp5_2020
ab_ssp5_2025 <- crops_2015_brazil - crops_ssp5_2025
ab_ssp5_2030 <- crops_2015_brazil - crops_ssp5_2030
ab_ssp5_2035 <- crops_2015_brazil - crops_ssp5_2035
ab_ssp5_2040 <- crops_2015_brazil - crops_ssp5_2040
ab_ssp5_2045 <- crops_2015_brazil - crops_ssp5_2045
ab_ssp5_2050 <- crops_2015_brazil - crops_ssp5_2050

```


Now, remove any pixels that change from cropland to urban in the future. Urbanization of cropland disqualifies it from potential restoration and is not useful for this project's analysis. This is done by subtracting projected urban areas (classified as value of 10) from the abandonment layers (classified value of 1). 
Values of 1 indicate cropland abandonment without transition to urban (1-0). Values of -9 indicate areas of abandonment that did transition to urban (1-10). 
```{r}
# SSP1 non-urbanized abandonment 2020-2050 --------------------
ab_ssp1_2020u <- ab_ssp1_2020 - urban_ssp1_2020
ab_ssp1_2025u <- ab_ssp1_2025 - urban_ssp1_2025
ab_ssp1_2030u <- ab_ssp1_2030 - urban_ssp1_2030
ab_ssp1_2035u <- ab_ssp1_2035 - urban_ssp1_2035
ab_ssp1_2040u <- ab_ssp1_2040 - urban_ssp1_2040
ab_ssp1_2045u <- ab_ssp1_2045 - urban_ssp1_2045
ab_ssp1_2050u <- ab_ssp1_2050 - urban_ssp1_2050

# SSP2 non-urbanized abandonment 2020-2050 --------------------
ab_ssp2_2020u <- ab_ssp2_2020 - urban_ssp2_2020
ab_ssp2_2025u <- ab_ssp2_2025 - urban_ssp2_2025
ab_ssp2_2030u <- ab_ssp2_2030 - urban_ssp2_2030
ab_ssp2_2035u <- ab_ssp2_2035 - urban_ssp2_2035
ab_ssp2_2040u <- ab_ssp2_2040 - urban_ssp2_2040
ab_ssp2_2045u <- ab_ssp2_2045 - urban_ssp2_2045
ab_ssp2_2050u <- ab_ssp2_2050 - urban_ssp2_2050

# SSP3 non-urbanized abandonment 2020-2050 --------------------
ab_ssp3_2020u <- ab_ssp3_2020 - urban_ssp3_2020
ab_ssp3_2025u <- ab_ssp3_2025 - urban_ssp3_2025
ab_ssp3_2030u <- ab_ssp3_2030 - urban_ssp3_2030
ab_ssp3_2035u <- ab_ssp3_2035 - urban_ssp3_2035
ab_ssp3_2040u <- ab_ssp3_2040 - urban_ssp3_2040
ab_ssp3_2045u <- ab_ssp3_2045 - urban_ssp3_2045
ab_ssp3_2050u <- ab_ssp3_2050 - urban_ssp3_2050

# SSP4 non-urbanized abandonment 2020-2050 --------------------
ab_ssp4_2020u <- ab_ssp4_2020 - urban_ssp4_2020
ab_ssp4_2025u <- ab_ssp4_2025 - urban_ssp4_2025
ab_ssp4_2030u <- ab_ssp4_2030 - urban_ssp4_2030
ab_ssp4_2035u <- ab_ssp4_2035 - urban_ssp4_2035
ab_ssp4_2040u <- ab_ssp4_2040 - urban_ssp4_2040
ab_ssp4_2045u <- ab_ssp4_2045 - urban_ssp4_2045
ab_ssp4_2050u <- ab_ssp4_2050 - urban_ssp4_2050

# SSP5 non-urbanized abandonment 2020-2050 --------------------
ab_ssp5_2020u <- ab_ssp5_2020 - urban_ssp5_2020
ab_ssp5_2025u <- ab_ssp5_2025 - urban_ssp5_2025
ab_ssp5_2030u <- ab_ssp5_2030 - urban_ssp5_2030
ab_ssp5_2035u <- ab_ssp5_2035 - urban_ssp5_2035
ab_ssp5_2040u <- ab_ssp5_2040 - urban_ssp5_2040
ab_ssp5_2045u <- ab_ssp5_2045 - urban_ssp5_2045
ab_ssp5_2050u <- ab_ssp5_2050 - urban_ssp5_2050
```

These layers may be useful for future study into the rates of LULC transition over time between cropland and urban classification. For this analysis, we want to isolate just pixels of non-urbanized cropland abandonment (value 1). In this next step, we'll reclassify the abandonment layers to only return pixels of non-urbanized abandonment (for ease of next analysis steps).

```{r}
# reclassify SSP1 non-urbanized cropland abandonment 2020-2050 
ab_ssp1_2020u[ab_ssp1_2020u != 1] <- 0  # only keeping values of 1
ab_ssp1_2025u[ab_ssp1_2025u != 1] <- 0
ab_ssp1_2030u[ab_ssp1_2030u != 1] <- 0
ab_ssp1_2035u[ab_ssp1_2035u != 1] <- 0
ab_ssp1_2040u[ab_ssp1_2040u != 1] <- 0
ab_ssp1_2045u[ab_ssp1_2045u != 1] <- 0
ab_ssp1_2050u[ab_ssp1_2050u != 1] <- 0

# reclassify SSP2 non-urbanized cropland abandonment 2020-2050 
ab_ssp2_2020u[ab_ssp2_2020u != 1] <- 0
ab_ssp2_2025u[ab_ssp2_2025u != 1] <- 0
ab_ssp2_2030u[ab_ssp2_2030u != 1] <- 0
ab_ssp2_2035u[ab_ssp2_2035u != 1] <- 0
ab_ssp2_2040u[ab_ssp2_2040u != 1] <- 0
ab_ssp2_2045u[ab_ssp2_2045u != 1] <- 0
ab_ssp2_2050u[ab_ssp2_2050u != 1] <- 0

# reclassify SSP3 non-urbanized cropland abandonment 2020-2050 
ab_ssp3_2020u[ab_ssp3_2020u != 1] <- 0
ab_ssp3_2025u[ab_ssp3_2025u != 1] <- 0
ab_ssp3_2030u[ab_ssp3_2030u != 1] <- 0
ab_ssp3_2035u[ab_ssp3_2035u != 1] <- 0
ab_ssp3_2040u[ab_ssp3_2040u != 1] <- 0
ab_ssp3_2045u[ab_ssp3_2045u != 1] <- 0
ab_ssp3_2050u[ab_ssp3_2050u != 1] <- 0

# reclassify SSP4 non-urbanized cropland abandonment 2020-2050 
ab_ssp4_2020u[ab_ssp4_2020u != 1] <- 0
ab_ssp4_2025u[ab_ssp4_2025u != 1] <- 0
ab_ssp4_2030u[ab_ssp4_2030u != 1] <- 0
ab_ssp4_2035u[ab_ssp4_2035u != 1] <- 0
ab_ssp4_2040u[ab_ssp4_2040u != 1] <- 0
ab_ssp4_2045u[ab_ssp4_2045u != 1] <- 0
ab_ssp4_2050u[ab_ssp4_2050u != 1] <- 0

# reclassify SSP5 non-urbanized cropland abandonment 2020-2050 
ab_ssp5_2020u[ab_ssp5_2020u != 1] <- 0
ab_ssp5_2025u[ab_ssp5_2025u != 1] <- 0
ab_ssp5_2030u[ab_ssp5_2030u != 1] <- 0
ab_ssp5_2035u[ab_ssp5_2035u != 1] <- 0
ab_ssp5_2040u[ab_ssp5_2040u != 1] <- 0
ab_ssp5_2045u[ab_ssp5_2045u != 1] <- 0
ab_ssp5_2050u[ab_ssp5_2050u != 1] <- 0
```

Then, values for an SSP will be summed to return one spatial raster indicating every pixel of cropland was projected to be abandoned *at some point* between 2020-2050. The number of years each pixel is abandoned may be useful for future analysis and addressing other research questions. For the sake of this analysis, we will reclassify summed values to all be 1. 

```{r}
# matrix for reclassification
reclass_df <- c(0, 0,
                1, 1, 
                2, 1, 
                3, 1,
                4, 1,
                5, 1,
                6, 1,
                7, 1)

reclass_mat <- matrix(reclass_df, ncol=2, byrow=TRUE)

# add SSP1 layers and reclassify
abandoned_ssp1 <- (ab_ssp1_2020u + ab_ssp1_2025u + ab_ssp1_2030u + ab_ssp1_2035u + 
                    ab_ssp1_2040u + ab_ssp1_2045u + ab_ssp1_2050u)
abandoned_ssp1 <- classify(abandoned_ssp1, reclass_mat)

# add SSP2 layers and reclassify
abandoned_ssp2 <- (ab_ssp2_2020u + ab_ssp2_2025u + ab_ssp2_2030u + ab_ssp2_2035u + 
                    ab_ssp2_2040u + ab_ssp2_2045u + ab_ssp2_2050u)
abandoned_ssp2 <- classify(abandoned_ssp2, reclass_mat)

# add SSP3 layers and reclassify
abandoned_ssp3 <- (ab_ssp3_2020u + ab_ssp3_2025u + ab_ssp3_2030u + 
                    ab_ssp3_2035u + ab_ssp3_2040u + ab_ssp3_2045u + ab_ssp3_2050u)
abandoned_ssp3 <- classify(abandoned_ssp3, reclass_mat)

# add SSP4 layers and reclassify
abandoned_ssp4 <- (ab_ssp4_2020u + ab_ssp4_2025u + ab_ssp4_2030u + 
                    ab_ssp4_2035u + ab_ssp4_2040u + ab_ssp4_2045u + ab_ssp4_2050u)
abandoned_ssp4 <- classify(abandoned_ssp4, reclass_mat)

# add SSP5 layers and reclassify
abandoned_ssp5 <- (ab_ssp5_2020u + ab_ssp5_2025u + ab_ssp5_2030u + 
                     ab_ssp5_2035u + ab_ssp5_2040u + ab_ssp5_2045u + ab_ssp5_2050u)
abandoned_ssp5 <- classify(abandoned_ssp5, reclass_mat)
```


## Export layers

Now we want to save these layers for future use in other scripts.
```{r}
# saves raster to local project in data folder
writeRaster(abandoned_ssp1, "data/outputs/SSP1_abandoned_cropland_brazil.tif")
writeRaster(abandoned_ssp2, "data/outputs/SSP2_abandoned_cropland_brazil.tif")
writeRaster(abandoned_ssp3, "data/outputs/SSP3_abandoned_cropland_brazil.tif")
writeRaster(abandoned_ssp4, "data/outputs/SSP4_abandoned_cropland_brazil.tif")
writeRaster(abandoned_ssp5, "data/outputs/SSP5_abandoned_cropland_brazil.tif")
```

