---
title: "abandoned_cropland_Brazil"
author: "Nick McManus"
date: '2022-10-06'
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: no
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)  #polygons
library(raster) #raster package
library(terra)  #newer/faster/better raster package
library(rnaturalearth)  #administrative boundaries data (DELTE)
library(geobr)  #administrative boundaries for Brazil
library(tidyverse) #always
```

## Overview
This analysis serves to find areas of projected cropland abandonment in Brazil between the years 2020 to 2050. Land use and land cover (LULC) data sourced from Chen et al. 2022. This dataset includes projected LULC under different shared socioeconomic pathways (SSPs) as well as a baseline LULC data from 2015. The outputs from this code — rasters containing every pixel of projected abandonment between 2020-2050 for a given SSP — will be utilized for further analysis into conservation optimization. More information on the source data can be found in the project README.md file. 

Source: Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6


## Read in data

First read in projected LULC for every five years between 2020-2050. There are LULC layers for each of the five different SSPs.

```{r}
# read in "baseline" 2015 LULC
PFT_2015_global <- rast("data/inputs/global_PFT_2015.tif")


# read in projected SSP1 LULC for 2020 - 2050 -------------------------------------------
ssp1_2020_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2020.tif")   #2020
ssp1_2025_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2025.tif")   #2025
ssp1_2030_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2030.tif")   #2030
ssp1_2035_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2035.tif")   #2035
ssp1_2040_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2040.tif")   #2040
ssp1_2045_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2045.tif")   #2045
ssp1_2050_global <- rast("data/inputs/SSP1_RCP26/global_PFT_SSP1_RCP26_2050.tif")   #2050


# read in projected SSP2 LULC for 2020 - 2050 -------------------------------------------
ssp2_2020_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2020.tif") 
ssp2_2025_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2025.tif")
ssp2_2030_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2030.tif") 
ssp2_2035_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2035.tif")
ssp2_2040_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2040.tif")
ssp2_2045_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2045.tif")
ssp2_2050_global <- rast("data/inputs/SSP2_RCP45/global_PFT_SSP2_RCP45_2050.tif")


# read in projected SSP3 LULC for 2020 - 2050 -------------------------------------------
ssp3_2020_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2020.tif")  
ssp3_2025_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2025.tif")  
ssp3_2030_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2030.tif")  
ssp3_2035_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2035.tif")  
ssp3_2040_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2040.tif")  
ssp3_2045_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2045.tif")  
ssp3_2050_global <- rast("data/inputs/SSP3_RCP70/global_PFT_SSP3_RCP70_2050.tif")  


# read in projected SSP4 LULC for 2020 - 2050 -------------------------------------------
ssp4_2020_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2020.tif") 
ssp4_2025_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2025.tif") 
ssp4_2030_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2030.tif") 
ssp4_2035_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2035.tif") 
ssp4_2040_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2040.tif") 
ssp4_2045_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2045.tif")
ssp4_2050_global <- rast("data/inputs/SSP4_RCP60/global_PFT_SSP4_RCP60_2050.tif") 


# read in projected SSP5 LULC for 2020 - 2050 -------------------------------------------
ssp5_2020_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2020.tif") 
ssp5_2025_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2025.tif")
ssp5_2030_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2030.tif")
ssp5_2035_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2035.tif")
ssp5_2040_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2040.tif")
ssp5_2045_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2045.tif")
ssp5_2050_global <- rast("data/inputs/SSP5_RCP85/global_PFT_SSP5_RCP85_2050.tif")
```


Brazil is the ROI for this analysis. Read in shape file for Brazil to later clip LULC layers. 
```{r}
# get brazil outline from geobr data
brazil <- read_country(year = 2020,         #most recent boundary data
                       simplified = FALSE)  #more accurate for spatial analyses
```


## Clipping and Masking 

Transform Brazil shapefile to match Chen LULC raster coordinate reference system. This is faster then re-projected a global raster to match Brazil's crs.
```{r}
# transform brazil to chen raster
brazil_trans <- brazil %>% 
  st_transform(crs = crs(PFT_2015_global)) %>% 
  vect() # make object a vector, b/c vects more agreeable to rasts
```

Now crop and mask each LULC raster to our `brazil_trans` vector.
- Note: the `terra::crop()` function that clips rasters is not to be confused with "crops" used in later variable names to indicate rasters reclassified to isolate cropland. 
```{r}
# crop baseline 2015 data
PFT_2015_brazil <- PFT_2015_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
# after clipping/masking, remove global object from environment
# to save on computing space
rm(PFT_2015_global)

#use this to transform back to original crs of brazil (EPSG4674)
#need to decide which projection is best for this analysis!
#terra::project(y = crs(brazil)) 


# crop and mask all SSP1 layers --------------------------
ssp1_2020_brazil <- ssp1_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp1_2020_global)

ssp1_2025_brazil <- ssp1_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp1_2025_global)

ssp1_2030_brazil <- ssp1_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp1_2030_global)

ssp1_2035_brazil <- ssp1_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp1_2035_global)

ssp1_2040_brazil <- ssp1_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp1_2040_global)

ssp1_2045_brazil <- ssp1_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp1_2045_global)

ssp1_2050_brazil <- ssp1_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp1_2050_global)


# crop and mask all SSP2 layers --------------------------
ssp2_2020_brazil <- ssp2_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp2_2020_global)

ssp2_2025_brazil <- ssp2_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp2_2025_global)

ssp2_2030_brazil <- ssp2_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp2_2030_global)

ssp2_2035_brazil <- ssp2_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp2_2035_global)

ssp2_2040_brazil <- ssp2_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp2_2040_global)

ssp2_2045_brazil <- ssp2_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp2_2045_global)

ssp2_2050_brazil <- ssp2_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp2_2050_global)


# crop and mask all SSP3 layers --------------------------
ssp3_2020_brazil <- ssp3_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp3_2020_global)

ssp3_2025_brazil <- ssp3_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp3_2025_global)

ssp3_2030_brazil <- ssp3_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp3_2030_global)

ssp3_2035_brazil <- ssp3_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp3_2035_global)

ssp3_2040_brazil <- ssp3_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp3_2040_global)

ssp3_2045_brazil <- ssp3_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp3_2045_global)

ssp3_2050_brazil <- ssp3_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp3_2050_global)

# crop and mask all SSP4 layers ---------------------------
ssp4_2020_brazil <- ssp4_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2020_global)

ssp4_2025_brazil <- ssp4_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2025_global)

ssp4_2030_brazil <- ssp4_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2030_global)

ssp4_2035_brazil <- ssp4_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2035_global)

ssp4_2040_brazil <- ssp4_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2040_global)

ssp4_2045_brazil <- ssp4_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp4_2045_global)

ssp4_2050_brazil <- ssp4_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp4_2050_global)


# crop and mask all SSP5 layers --------------------------
ssp5_2020_brazil <- ssp5_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp5_2020_global)

ssp5_2025_brazil <- ssp5_2025_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp5_2025_global)

ssp5_2030_brazil <- ssp5_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp5_2030_global)

ssp5_2035_brazil <- ssp5_2035_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp5_2035_global)

ssp5_2040_brazil <- ssp5_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp5_2040_global)

ssp5_2045_brazil <- ssp5_2045_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) 
rm(ssp5_2045_global)

ssp5_2050_brazil <- ssp5_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans)
rm(ssp5_2050_global)
```


## Reclassify

First, reclassify each LULC layer to isolate pixels of cropland (value 18) and urban (value 19) from the raster. Each layer will be reclassified to make all other variables 0.

```{r}
# create matrix for reclassification
reclass_df <- c(1, 0,	  #Water
                2, 0,	  #Broadleaf evergreen tree, tropical
                3, 0,	  #Broadleaf evergreen tree, temperate
                4, 0,	  #Broadleaf deciduous tree, tropical
                5, 0,	  #Broadleaf deciduous tree, temperate
                6, 0,	  #Broadleaf deciduous tree, boreal
                7, 0,	  #Needleleaf evergreen tree, temperate
                8, 0,	  #Needleleaf evergreen tree, boreal
                9, 0,	  #Needleleaf deciduous tree
                10, 0,	#Broadleaf evergreen shrub, temperate
                11, 0,	#Broadleaf deciduous shrub, temperate
                12, 0,	#Broadleaf deciduous shrub, boreal
                13, 0,	#C3 grass, arctic
                14, 0,	#C3 grass
                15, 0,	#C4 grass
                16, 0,	#Mixed C3/C4 grass
                17, 0,	#Barren
                18, 1,	#Cropland
                19, 10,	#Urban
                20, 0,  #Permanent snow and ice
                128, 0)	#N/A values

reclass_m <- matrix(reclass_df, ncol=2, byrow=TRUE)

# baseline 2015 data -------------------------------------------
crops_urban_2015_brazil <- classify(PFT_2015_brazil, reclass_m)
  
  
# SSP1 layers for 2020 - 2050 ----------------------------------
ssp1_2020_crops_urban <- classify(ssp1_2020_brazil, reclass_m)
ssp1_2025_crops_urban <- classify(ssp1_2025_brazil, reclass_m)
ssp1_2030_crops_urban <- classify(ssp1_2030_brazil, reclass_m)
ssp1_2035_crops_urban <- classify(ssp1_2035_brazil, reclass_m)
ssp1_2040_crops_urban <- classify(ssp1_2040_brazil, reclass_m)
ssp1_2045_crops_urban <- classify(ssp1_2045_brazil, reclass_m)
ssp1_2050_crops_urban <- classify(ssp1_2050_brazil, reclass_m)

# remove old raster to save environment space (optional)
rm(ssp1_2020_brazil)
rm(ssp1_2025_brazil)
rm(ssp1_2030_brazil)
rm(ssp1_2035_brazil)
rm(ssp1_2040_brazil)
rm(ssp1_2045_brazil)
rm(ssp1_2050_brazil)

# SSP2 layers for 2020 - 2050 -----------------------------------
ssp2_2020_crops_urban <- classify(ssp2_2020_brazil, reclass_m)
ssp2_2025_crops_urban <- classify(ssp2_2025_brazil, reclass_m)
ssp2_2030_crops_urban <- classify(ssp2_2030_brazil, reclass_m)
ssp2_2035_crops_urban <- classify(ssp2_2035_brazil, reclass_m)
ssp2_2040_crops_urban <- classify(ssp2_2040_brazil, reclass_m)
ssp2_2045_crops_urban <- classify(ssp2_2045_brazil, reclass_m)
ssp2_2050_crops_urban <- classify(ssp2_2050_brazil, reclass_m)

rm(ssp2_2020_brazil)
rm(ssp2_2025_brazil)
rm(ssp2_2030_brazil)
rm(ssp2_2035_brazil)
rm(ssp2_2040_brazil)
rm(ssp2_2045_brazil)
rm(ssp2_2050_brazil)

# SSP3 layers for 2020 - 2050 -----------------------------------
ssp3_2020_crops_urban <- classify(ssp3_2020_brazil, reclass_m)
ssp3_2025_crops_urban <- classify(ssp3_2025_brazil, reclass_m)
ssp3_2030_crops_urban <- classify(ssp3_2030_brazil, reclass_m)
ssp3_2035_crops_urban <- classify(ssp3_2035_brazil, reclass_m)
ssp3_2040_crops_urban <- classify(ssp3_2040_brazil, reclass_m)
ssp3_2045_crops_urban <- classify(ssp3_2045_brazil, reclass_m)
ssp3_2050_crops_urban <- classify(ssp3_2050_brazil, reclass_m)
  
rm(ssp3_2020_brazil)
rm(ssp3_2025_brazil)
rm(ssp3_2030_brazil)
rm(ssp3_2035_brazil)
rm(ssp3_2040_brazil)
rm(ssp3_2045_brazil)
rm(ssp3_2050_brazil)

# SSP4 layers for 2020 - 2050 -----------------------------------
ssp4_2020_crops_urban <- classify(ssp4_2020_brazil, reclass_m)
ssp4_2025_crops_urban <- classify(ssp4_2025_brazil, reclass_m)
ssp4_2030_crops_urban <- classify(ssp4_2030_brazil, reclass_m)
ssp4_2035_crops_urban <- classify(ssp4_2035_brazil, reclass_m)
ssp4_2040_crops_urban <- classify(ssp4_2040_brazil, reclass_m)
ssp4_2045_crops_urban <- classify(ssp4_2045_brazil, reclass_m)
ssp4_2050_crops_urban <- classify(ssp4_2050_brazil, reclass_m)

rm(ssp4_2020_brazil)
rm(ssp4_2025_brazil)
rm(ssp4_2030_brazil)
rm(ssp4_2035_brazil)
rm(ssp4_2040_brazil)
rm(ssp4_2045_brazil)
rm(ssp4_2050_brazil)
  
# SSP5 layers for 2020 - 2050 -----------------------------------
ssp5_2020_crops_urban <- classify(ssp5_2020_brazil, reclass_m)
ssp5_2025_crops_urban <- classify(ssp5_2025_brazil, reclass_m)
ssp5_2030_crops_urban <- classify(ssp5_2030_brazil, reclass_m)
ssp5_2035_crops_urban <- classify(ssp5_2035_brazil, reclass_m)
ssp5_2040_crops_urban <- classify(ssp5_2040_brazil, reclass_m)
ssp5_2045_crops_urban <- classify(ssp5_2045_brazil, reclass_m)
ssp5_2050_crops_urban <- classify(ssp5_2050_brazil, reclass_m)

rm(ssp5_2020_brazil)
rm(ssp5_2025_brazil)
rm(ssp5_2030_brazil)
rm(ssp5_2035_brazil)
rm(ssp5_2040_brazil)
rm(ssp5_2045_brazil)
rm(ssp5_2050_brazil)
```


## Calculate abandonment

Now that we've isolated Brazilian cropland, we want to determine where current cropland is projected to become "abandoned" in the future. This will be done with a simple raster calculation, subtracting the future cropland raster from the current cropland raster. After the subtraction, values of 0 indicate land that either remained cropland (1-1), remained urban (10-10), or remained non-cropland (0-0). Values of -1 indicate new future cropland (0-1), values of -10 indicate areas of new urban (0-10), and values of -9 represent cropland that became urbanized (1-10). Values of *1* indicate abandoned cropland (1-0) that did not become urban, which are our pixels of interest!

This calculation will be repeated for each projected raster (2020-2050) against the baseline raster (2015). 

```{r}
# SSP1 abandonment 2020-2050 -----------------------------------------
ssp1_2020_abandon <- crops_urban_2015_brazil - ssp1_2020_crops_urban
ssp1_2025_abandon <- crops_urban_2015_brazil - ssp1_2025_crops_urban
ssp1_2030_abandon <- crops_urban_2015_brazil - ssp1_2030_crops_urban
ssp1_2035_abandon <- crops_urban_2015_brazil - ssp1_2035_crops_urban
ssp1_2040_abandon <- crops_urban_2015_brazil - ssp1_2040_crops_urban
ssp1_2045_abandon <- crops_urban_2015_brazil - ssp1_2045_crops_urban
ssp1_2050_abandon <- crops_urban_2015_brazil - ssp1_2050_crops_urban

# SSP2 abandonment 2020-2050 -----------------------------------------
ssp2_2020_abandon <- crops_urban_2015_brazil - ssp2_2020_crops_urban
ssp2_2025_abandon <- crops_urban_2015_brazil - ssp2_2025_crops_urban
ssp2_2030_abandon <- crops_urban_2015_brazil - ssp2_2030_crops_urban
ssp2_2035_abandon <- crops_urban_2015_brazil - ssp2_2035_crops_urban
ssp2_2040_abandon <- crops_urban_2015_brazil - ssp2_2040_crops_urban
ssp2_2045_abandon <- crops_urban_2015_brazil - ssp2_2045_crops_urban
ssp2_2050_abandon <- crops_urban_2015_brazil - ssp2_2050_crops_urban


# SSP3 abandonment 2020-2050 -----------------------------------------
ssp3_2020_abandon <- crops_urban_2015_brazil - ssp3_2020_crops_urban
ssp3_2025_abandon <- crops_urban_2015_brazil - ssp3_2025_crops_urban
ssp3_2030_abandon <- crops_urban_2015_brazil - ssp3_2030_crops_urban
ssp3_2035_abandon <- crops_urban_2015_brazil - ssp3_2035_crops_urban
ssp3_2040_abandon <- crops_urban_2015_brazil - ssp3_2040_crops_urban
ssp3_2045_abandon <- crops_urban_2015_brazil - ssp3_2045_crops_urban
ssp3_2050_abandon <- crops_urban_2015_brazil - ssp3_2050_crops_urban


# SSP4 abandonment 2020-2050 -----------------------------------------
ssp4_2020_abandon <- crops_urban_2015_brazil - ssp4_2020_crops_urban
ssp4_2025_abandon <- crops_urban_2015_brazil - ssp4_2025_crops_urban
ssp4_2030_abandon <- crops_urban_2015_brazil - ssp4_2030_crops_urban
ssp4_2035_abandon <- crops_urban_2015_brazil - ssp4_2035_crops_urban
ssp4_2040_abandon <- crops_urban_2015_brazil - ssp4_2040_crops_urban
ssp4_2045_abandon <- crops_urban_2015_brazil - ssp4_2045_crops_urban
ssp4_2050_abandon <- crops_urban_2015_brazil - ssp4_2050_crops_urban


# SSP5 abandonment 2020-2050 -----------------------------------------
ssp5_2020_abandon <- crops_urban_2015_brazil - ssp5_2020_crops_urban
ssp5_2025_abandon <- crops_urban_2015_brazil - ssp5_2025_crops_urban
ssp5_2030_abandon <- crops_urban_2015_brazil - ssp5_2030_crops_urban
ssp5_2035_abandon <- crops_urban_2015_brazil - ssp5_2035_crops_urban
ssp5_2040_abandon <- crops_urban_2015_brazil - ssp5_2040_crops_urban
ssp5_2045_abandon <- crops_urban_2015_brazil - ssp5_2045_crops_urban
ssp5_2050_abandon <- crops_urban_2015_brazil - ssp5_2050_crops_urban

```

These layers may be useful for future study into the rates of LULC transition over time between cropland and urban classification. For this analysis, we want to isolate just pixels of non-urbanized cropland abandonment (*value 1*). In this next step, we'll reclassify the abandonment layers to only return pixels of non-urbanized abandonment.

```{r}
# reclassify SSP1 non-urbanized cropland abandonment 2020-2050 
## Convert all values that aren't 1 to 0. Use 0 rather than NA
## so rasters can be added up correctly in next step
ssp1_2020_abandon[ssp1_2020_abandon != 1] <- 0
ssp1_2025_abandon[ssp1_2025_abandon != 1] <- 0
ssp1_2030_abandon[ssp1_2030_abandon != 1] <- 0
ssp1_2035_abandon[ssp1_2035_abandon != 1] <- 0
ssp1_2040_abandon[ssp1_2040_abandon != 1] <- 0
ssp1_2045_abandon[ssp1_2045_abandon != 1] <- 0
ssp1_2050_abandon[ssp1_2050_abandon != 1] <- 0

# reclassify SSP2 non-urbanized cropland abandonment 2020-2050 
ssp2_2020_abandon[ssp2_2020_abandon != 1] <- 0  
ssp2_2025_abandon[ssp2_2025_abandon != 1] <- 0
ssp2_2030_abandon[ssp2_2030_abandon != 1] <- 0
ssp2_2035_abandon[ssp2_2035_abandon != 1] <- 0
ssp2_2040_abandon[ssp2_2040_abandon != 1] <- 0
ssp2_2045_abandon[ssp2_2045_abandon != 1] <- 0
ssp2_2050_abandon[ssp2_2050_abandon != 1] <- 0

# reclassify SSP3 non-urbanized cropland abandonment 2020-2050 
ssp3_2020_abandon[ssp3_2020_abandon != 1] <- 0 
ssp3_2025_abandon[ssp3_2025_abandon != 1] <- 0
ssp3_2030_abandon[ssp3_2030_abandon != 1] <- 0
ssp3_2035_abandon[ssp3_2035_abandon != 1] <- 0
ssp3_2040_abandon[ssp3_2040_abandon != 1] <- 0
ssp3_2045_abandon[ssp3_2045_abandon != 1] <- 0
ssp3_2050_abandon[ssp3_2050_abandon != 1] <- 0

# reclassify SSP4 non-urbanized cropland abandonment 2020-2050 
ssp4_2020_abandon[ssp4_2020_abandon != 1] <- 0  
ssp4_2025_abandon[ssp4_2025_abandon != 1] <- 0
ssp4_2030_abandon[ssp4_2030_abandon != 1] <- 0
ssp4_2035_abandon[ssp4_2035_abandon != 1] <- 0
ssp4_2040_abandon[ssp4_2040_abandon != 1] <- 0
ssp4_2045_abandon[ssp4_2045_abandon != 1] <- 0
ssp4_2050_abandon[ssp4_2050_abandon != 1] <- 0

# reclassify SSP5 non-urbanized cropland abandonment 2020-2050 
ssp5_2020_abandon[ssp5_2020_abandon != 1] <- 0  
ssp5_2025_abandon[ssp5_2025_abandon != 1] <- 0
ssp5_2030_abandon[ssp5_2030_abandon != 1] <- 0
ssp5_2035_abandon[ssp5_2035_abandon != 1] <- 0
ssp5_2040_abandon[ssp5_2040_abandon != 1] <- 0
ssp5_2045_abandon[ssp5_2045_abandon != 1] <- 0
ssp5_2050_abandon[ssp5_2050_abandon != 1] <- 0
```

Then, values for an SSP will be summed to return one spatial raster indicating every pixel of cropland was projected to be abandoned *at some point* between 2020-2050. The number of years each pixel is abandoned may be useful for future analysis and addressing other research questions. This could potentially be used as a feature when formulating the prioritization problem. Values of 0 represent land not available for restoration and will be removed from raster.

```{r}
# add SSP1 layers and reclassify
abandoned_ssp1 <- (ssp1_2020_abandon + ssp1_2025_abandon + 
                     ssp1_2030_abandon + ssp1_2035_abandon + 
                     ssp1_2040_abandon + ssp1_2045_abandon + 
                     ssp1_2050_abandon)

# replace all 0 values with N/A
abandoned_ssp1[abandoned_ssp1 == 0] <- NA 



# add SSP2 layers and reclassify ------------------------------------
abandoned_ssp2 <- (ssp2_2020_abandon + ssp2_2025_abandon +
                     ssp2_2030_abandon + ssp2_2035_abandon +
                     ssp2_2040_abandon + ssp2_2045_abandon +
                     ssp2_2050_abandon)

abandoned_ssp2[abandoned_ssp2 == 0] <- NA

# add SSP3 layers and reclassify ------------------------------------
abandoned_ssp3 <- (ssp3_2020_abandon + ssp3_2025_abandon +
                     ssp3_2030_abandon + ssp3_2035_abandon +
                     ssp3_2040_abandon + ssp3_2045_abandon +
                     ssp3_2050_abandon)

abandoned_ssp3[abandoned_ssp3 == 0] <- NA

# add SSP4 layers and reclassify ------------------------------------
abandoned_ssp4 <- (ssp4_2020_abandon + ssp4_2025_abandon +
                     ssp4_2030_abandon + ssp4_2035_abandon +
                     ssp4_2040_abandon + ssp4_2045_abandon +
                     ssp4_2050_abandon)

abandoned_ssp4[abandoned_ssp4 == 0] <- NA

# add SSP5 layers and reclassify ------------------------------------
abandoned_ssp5 <- (ssp5_2020_abandon + ssp5_2025_abandon +
                     ssp5_2030_abandon + ssp5_2035_abandon +
                     ssp5_2040_abandon + ssp5_2045_abandon +
                     ssp5_2050_abandon)

abandoned_ssp5[abandoned_ssp5 == 0] <- NA
```

```{r}
# Use this code chunk to change final layer CRS if needed....
#abandoned_ssp1 <- project(abandoned_ssp1, "epsg:4326") 
#abandoned_ssp2 <- project(abandoned_ssp2, "epsg:4326")
#abandoned_ssp3 <- project(abandoned_ssp3, "epsg:4326")
#abandoned_ssp4 <- project(abandoned_ssp4, "epsg:4326")
#abandoned_ssp5 <- project(abandoned_ssp5, "epsg:4326")
```

## Export layers

Now we want to save these layers for future use in other scripts.
```{r}
# saves raster to local project in data folder
writeRaster(abandoned_ssp1, "data/outputs/SSP1_abandoned_cropland_brazil.tif", overwrite = TRUE)
writeRaster(abandoned_ssp2, "data/outputs/SSP2_abandoned_cropland_brazil.tif", overwrite = TRUE)
writeRaster(abandoned_ssp3, "data/outputs/SSP3_abandoned_cropland_brazil.tif", overwrite = TRUE)
writeRaster(abandoned_ssp4, "data/outputs/SSP4_abandoned_cropland_brazil.tif", overwrite = TRUE)
writeRaster(abandoned_ssp5, "data/outputs/SSP5_abandoned_cropland_brazil.tif", overwrite = TRUE)
```

