---
title: "abandoned_cropland"
author: "Nick McManus"
date: '2022-10-06'
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)  #polygons
library(raster) #raster package
library(terra)  #newer/faster/better raster package
library(rnaturalearth)  #administrative boundaries data
library(exactextractr)  #zonal stats
library(kableExtra) #tables
library(tidyverse) #always
library(caret) # for confusion matrix
```

## Overview
First read in projected LULC from Chen et al. 2022. Then, crop and mask this global projection to just Brazil. Finally, reclassify the raster to keep only pixels representing cropland.

Source: Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6


## Read in the data

First read in projected LULC from Chen et al. for each decade between 2020-2050.There are LULC layers for each of the five different SSPs.

```{r}
# read in "baseline" 2015 LULC
PFT_2015_global <- rast("data/global_PFT_2015.tif")


# read in projected SSP1 LULC for 2020 - 2050
ssp1_2020_global <- rast("data/SSP1_RCP26/global_PFT_SSP1_RCP26_2020.tif")   #2020
ssp1_2030_global <- rast("data/SSP1_RCP26/global_PFT_SSP1_RCP26_2030.tif")   #2030
ssp1_2040_global <- rast("data/SSP1_RCP26/global_PFT_SSP1_RCP26_2040.tif")   #2040
ssp1_2050_global <- rast("data/SSP1_RCP26/global_PFT_SSP1_RCP26_2050.tif")   #2050

# read in projected SSP2 LULC for 2020 - 2050
ssp2_2020_global <- rast("data/SSP2_RCP45/global_PFT_SSP2_RCP45_2020.tif")   #2020
ssp2_2030_global <- rast("data/SSP2_RCP45/global_PFT_SSP2_RCP45_2030.tif")   #2030
ssp2_2040_global <- rast("data/SSP2_RCP45/global_PFT_SSP2_RCP45_2040.tif")   #2040
ssp2_2050_global <- rast("data/SSP2_RCP45/global_PFT_SSP2_RCP45_2050.tif")   #2050

# read in projected SSP3 LULC for 2020 - 2050
ssp3_2020_global <- rast("data/SSP3_RCP70/global_PFT_SSP3_RCP70_2020.tif")   #2020
ssp3_2030_global <- rast("data/SSP3_RCP70/global_PFT_SSP3_RCP70_2030.tif")   #2030
ssp3_2040_global <- rast("data/SSP3_RCP70/global_PFT_SSP3_RCP70_2040.tif")   #2040
ssp3_2050_global <- rast("data/SSP3_RCP70/global_PFT_SSP3_RCP70_2050.tif")   #2050

# read in projected SSP4 LULC for 2020 - 2050
ssp4_2020_global <- rast("data/SSP4_RCP60/global_PFT_SSP4_RCP60_2020.tif")   #2020
ssp4_2020_global <- rast("data/SSP4_RCP60/global_PFT_SSP4_RCP60_2030.tif")   #2030
ssp4_2020_global <- rast("data/SSP4_RCP60/global_PFT_SSP4_RCP60_2040.tif")   #2040
ssp4_2020_global <- rast("data/SSP4_RCP60/global_PFT_SSP4_RCP60_2050.tif")   #2050

# read in projected SSP5 LULC for 2020 - 2050
ssp5_2020_global <- rast("data/SSP5_RCP85/global_PFT_SSP5_RCP85_2020.tif")   #2020
ssp5_2030_global <- rast("data/SSP5_RCP85/global_PFT_SSP5_RCP85_2030.tif")   #2030
ssp5_2040_global <- rast("data/SSP5_RCP85/global_PFT_SSP5_RCP85_2040.tif")   #2040
ssp5_2050_global <- rast("data/SSP5_RCP85/global_PFT_SSP5_RCP85_2050.tif")   #2050
```


Brazil is the ROI for this analysis. Read in shape file for Brazil to later clip LULC layers. 
```{r}
# get brazil from rnaturalearth data
brazil <- ne_countries(
  scale = "medium",
  country = "Brazil",
  returnclass = "sf") %>% 
  dplyr::select(sovereignt)

# plot to check
ggplot() +
  geom_sf(data = brazil)
```

## Clipping and Masking 

Transform Brazil shapefile to match Chen LULC raster coordinate reference system. This is faster then re-projected a global raster to match Brazil's crs.
```{r}
# transform brazil to chen raster
brazil_trans <- brazil %>% 
  st_transform(crs = crs(PFT_2015_global)) %>% 
  vect() # make object a vector, b/c vects more agreeable to rasts
```

Now crop and mask each LULC raster to our `brazil_trans` vector. Then, project raster back to our desired crs.
- Note: terra::crop() function is not to be confused with "cropland" 
```{r}
# crop baseline 2015 data
PFT_2015_brazil <- PFT_2015_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP1 layers ----------------
ssp1_2020_brazil <- ssp1_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2030_brazil <- ssp1_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2040_brazil <- ssp1_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp1_2050_brazil <- ssp1_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP2 layers ----------------
ssp2_2020_brazil <- ssp2_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2030_brazil <- ssp2_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2040_brazil <- ssp2_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp2_2050_brazil <- ssp2_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP3 layers ----------------
ssp3_2020_brazil <- ssp3_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2030_brazil <- ssp3_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2040_brazil <- ssp3_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp3_2050_brazil <- ssp3_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP4 layers -----------------
ssp4_2020_brazil <- ssp4_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2030_brazil <- ssp4_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2040_brazil <- ssp4_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp4_2050_brazil <- ssp4_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))


# crop and mask all SSP5 layers ----------------
ssp5_2020_brazil <- ssp5_2020_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2030_brazil <- ssp5_2030_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2040_brazil <- ssp5_2040_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))

ssp5_2050_brazil <- ssp5_2050_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  terra::project(y = crs(brazil))
```


## Reclassify

Reclassify each LULC layer to isolate cropland (value 18) from the raster. 

```{r}
# baseline 2015 data
crops_2015_brazil <- PFT_2015_brazil 
  crops_2015_brazil[crops_2015_brazil != 18] <- 0 # if value isn't 18, make 0
  crops_2015_brazil[crops_2015_brazil == 18] <- 1 # if value is 18, make 1
  
  
# SSP1 layers for 2020 - 2050 -------------------
crops_ssp1_2020 <- ssp1_2020_brazil
  crops_ssp1_2020[crops_ssp1_2020 != 18] <- 0
  crops_ssp1_2020[crops_ssp1_2020 == 18] <- 1

crops_ssp1_2030 <- ssp1_2030_brazil   
  crops_ssp1_2030[crops_ssp1_2030 != 18] <- 0
  crops_ssp1_2030[crops_ssp1_2030 == 18] <- 1
  
crops_ssp1_2040 <- ssp1_2040_brazil
  crops_ssp1_2040[crops_ssp1_2040 != 18] <- 0
  crops_ssp1_2040[crops_ssp1_2040 == 18] <- 1
  
crops_ssp1_2050 <- ssp1_2050_brazil
  crops_ssp1_2050[crops_ssp1_2050 != 18] <- 0
  crops_ssp1_2050[crops_ssp1_2050 == 18] <- 1
  

# SSP2 layers for 2020 - 2050 -------------------
crops_ssp2_2020 <- ssp2_2020_brazil
  crops_ssp2_2020[crops_ssp2_2020 != 18] <- 0
  crops_ssp2_2020[crops_ssp2_2020 == 18] <- 1

crops_ssp2_2030 <- ssp2_2030_brazil   
  crops_ssp2_2030[crops_ssp2_2030 != 18] <- 0
  crops_ssp2_2030[crops_ssp2_2030 == 18] <- 1
  
crops_ssp2_2040 <- ssp2_2040_brazil
  crops_ssp2_2040[crops_ssp2_2040 != 18] <- 0
  crops_ssp2_2040[crops_ssp2_2040 == 18] <- 1
  
crops_ssp2_2050 <- ssp2_2050_brazil
  crops_ssp2_2050[crops_ssp2_2050 != 18] <- 0
  crops_ssp2_2050[crops_ssp2_2050 == 18] <- 1

  
# SSP3 layers for 2020 - 2050 -------------------
crops_ssp3_2020 <- ssp3_2020_brazil
  crops_ssp3_2020[crops_ssp3_2020 != 18] <- 0
  crops_ssp3_2020[crops_ssp3_2020 == 18] <- 1

crops_ssp3_2030 <- ssp3_2030_brazil   
  crops_ssp3_2030[crops_ssp3_2030 != 18] <- 0
  crops_ssp3_2030[crops_ssp3_2030 == 18] <- 1
  
crops_ssp3_2040 <- ssp3_2040_brazil
  crops_ssp3_2040[crops_ssp3_2040 != 18] <- 0
  crops_ssp3_2040[crops_ssp3_2040 == 18] <- 1
  
crops_ssp3_2050 <- ssp3_2050_brazil
  crops_ssp3_2050[crops_ssp3_2050 != 18] <- 0
  crops_ssp3_2050[crops_ssp3_2050 == 18] <- 1
  

# SSP4 layers for 2020 - 2050 -------------------
crops_ssp4_2020 <- ssp4_2020_brazil
  crops_ssp4_2020[crops_ssp4_2020 != 18] <- 0
  crops_ssp4_2020[crops_ssp4_2020 == 18] <- 1

crops_ssp4_2030 <- ssp4_2030_brazil   
  crops_ssp4_2030[crops_ssp4_2030 != 18] <- 0
  crops_ssp4_2030[crops_ssp4_2030 == 18] <- 1
  
crops_ssp4_2040 <- ssp4_2040_brazil
  crops_ssp4_2040[crops_ssp4_2040 != 18] <- 0
  crops_ssp4_2040[crops_ssp4_2040 == 18] <- 1
  
crops_ssp4_2050 <- ssp4_2050_brazil
  crops_ssp4_2050[crops_ssp4_2050 != 18] <- 0
  crops_ssp4_2050[crops_ssp4_2050 == 18] <- 1
  
  
# SSP5 layers for 2020 - 2050 -------------------
crops_ssp5_2020 <- ssp5_2020_brazil
  crops_ssp5_2020[crops_ssp5_2020 != 18] <- 0
  crops_ssp5_2020[crops_ssp5_2020 == 18] <- 1

crops_ssp5_2030 <- ssp5_2030_brazil   
  crops_ssp5_2030[crops_ssp5_2030 != 18] <- 0
  crops_ssp5_2030[crops_ssp5_2030 == 18] <- 1
  
crops_ssp5_2040 <- ssp5_2040_brazil
  crops_ssp5_2040[crops_ssp5_2040 != 18] <- 0
  crops_ssp5_2040[crops_ssp5_2040 == 18] <- 1
  
crops_ssp5_2050 <- ssp5_2050_brazil
  crops_ssp5_2050[crops_ssp5_2050 != 18] <- 0
  crops_ssp5_2050[crops_ssp5_2050 == 18] <- 1

```


## Calculate abandonment

Now that we've isolated Brazilian cropland, we want to determine where current cropland is projected to become "abandoned" in the future. This will done with a simple raster calculation, subtracting the future cropland raster from the current cropland raster. Values returning X will 

```{r}
# find abandoned pixels

# cropland abandonment for SSP1 for 2020 - 2050
abandoned_SSP1_2020 <- crops_ssp1_2020 - crops_2015_brazil

plot(abandoned_SSP1_2020)
```




